---
title: "TF-IDF: Finding Distinctive Vocabulary"
engine: knitr
format:
  html:
    fig-width: 10
    fig-height: 12
    dpi: 300
editor_options: 
  chunk_output_type: inline
---

```{r}
#| include: false
# This is just to render the document correctly in the CI/CD pipeline
library(tidyverse)
library(tidytext)

comments <- readr::read_csv("../../data/clean/comments_preprocessed.csv") 
```

So far, we have explored word frequencies and n-grams to understand common terms and phrases in our text data. However, simply counting words has a limitation: some words are frequent because they appear often across *all* documents, not because they are particularly meaningful for a specific document or group.

For example, in our Severance dataset, words like "season," "episode," and "show" might appear frequently in comments about both Season 1 and Season 2. While these words are common, they don't help us understand what makes each season's discussion *distinctive*.

This is where **TF-IDF (Term Frequency-Inverse Document Frequency)** becomes useful. TF-IDF is a statistical measure that evaluates how important a word is to a document in a collection of documents (corpus). It helps us identify words that are frequent in one document but rare across the entire corpus—precisely the words that make a document unique.

## Understanding TF-IDF

TF-IDF combines two metrics:

1. **Term Frequency (TF)**: How often a word appears in a document
2. **Inverse Document Frequency (IDF)**: How rare a word is across all documents

The formula is:

$$\text{TF-IDF} = \text{TF} \times \text{IDF}$$

Where:

$$\text{TF}(t, d) = \frac{\text{count of term } t \text{ in document } d}{\text{total terms in document } d}$$

$$\text{IDF}(t) = \log\left(\frac{\text{total number of documents}}{\text{number of documents containing term } t}\right)$$

A word gets a **high TF-IDF score** when:
- It appears frequently in a particular document (high TF)
- It appears in few other documents (high IDF)

A word gets a **low TF-IDF score** when:
- It appears in many documents (low IDF), even if it's frequent in one document

::: {.callout-note title="Why logarithm in IDF?" collapse="true"}
The logarithm in the IDF calculation dampens the effect of very rare words. Without it, a word that appears in only one document would dominate the scores. The log transformation ensures that the IDF increases more slowly as words become rarer, creating a more balanced weighting system.
:::

## Calculating TF-IDF

In our case, we want to compare the vocabulary between Season 1 and Season 2 comments. We'll treat each season as a "document" and calculate TF-IDF to find which words are distinctive to each season.

First, we need to extract season information from the `id` column and tokenize the comments:

```{r}
# Calculate TF-IDF by season
comments_tfidf <- comments %>%
  mutate(season = str_extract(id, "s[12]")) %>%  # Extract season (s1 or s2)
  unnest_tokens(word, comments) %>%               # Tokenize into words
  count(season, word, sort = TRUE)                # Count words per season

head(comments_tfidf)
```

Now we can apply the `bind_tf_idf()` function from the `tidytext` package, which automatically calculates TF, IDF, and TF-IDF for us:

```{r}
# Apply TF-IDF calculation
comments_tfidf <- comments_tfidf %>%
  bind_tf_idf(word, season, n)

head(comments_tfidf, 15)
```

The resulting data frame includes:
- `tf`: Term frequency (proportion of times the word appears in that season)
- `idf`: Inverse document frequency (how rare the word is across seasons)
- `tf_idf`: The product of TF and IDF

Let's examine the top words by TF-IDF for each season:

```{r}
# Top 10 distinctive words per season
comments_tfidf %>%
  group_by(season) %>%
  slice_max(tf_idf, n = 10)
```

Notice how these words are much more specific and meaningful than simply looking at the most frequent words. These are the words that truly characterize each season's discussion.

## Visualizing Distinctive Vocabulary

To better understand the distinctive vocabulary of each season, we can create a visualization comparing the top TF-IDF words:

```{r}
# Prepare data for visualization
top_tfidf_words <- comments_tfidf %>%
  group_by(season) %>%
  slice_max(tf_idf, n = 15) %>%
  ungroup() %>%
  mutate(word = reorder_within(word, tf_idf, season))

# Plot distinctive vocabulary by season
ggplot(top_tfidf_words, aes(tf_idf, word, fill = season)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~season, scales = "free") +
  scale_y_reordered() +
  labs(
    x = "TF-IDF",
    y = NULL,
    title = "Distinctive Vocabulary by Season"
  ) +
  theme_minimal()
```

This visualization clearly shows which words are most characteristic of each season's discussions. Words with higher TF-IDF scores are those that appear frequently in one season but not in the other, making them useful markers of distinctive content.

## Comparing TF-IDF to Raw Frequency

To appreciate the value of TF-IDF, let's compare it to simple word counts. We'll look at the top words by frequency versus the top words by TF-IDF for Season 1:

```{r}
# Top words by raw frequency for Season 1
top_freq_s1 <- comments %>%
  filter(grepl("^s1", id)) %>%
  unnest_tokens(word, comments) %>%
  count(word, sort = TRUE) %>%
  head(15)

# Top words by TF-IDF for Season 1
top_tfidf_s1 <- comments_tfidf %>%
  filter(season == "s1") %>%
  arrange(desc(tf_idf)) %>%
  head(15)

# Compare
cat("=== Top 15 words by frequency (Season 1) ===\n")
print(top_freq_s1)

cat("\n=== Top 15 words by TF-IDF (Season 1) ===\n")
print(top_tfidf_s1 %>% select(word, n, tf_idf))
```

The raw frequency list likely includes many words that are common across both seasons, while the TF-IDF list highlights words that are specifically important to Season 1 discussions.

## When to Use TF-IDF

TF-IDF is particularly useful for:

1. **Document comparison**: Identifying what makes each document unique in a collection
2. **Feature extraction**: Preparing text data for machine learning by emphasizing distinctive words
3. **Topic discovery**: Finding characteristic vocabulary for different groups or categories
4. **Search and retrieval**: Ranking documents by relevance to a query (search engines use variations of TF-IDF)

::: {.callout-tip title="Limitations of TF-IDF"}
While TF-IDF is powerful, it has some limitations:

- **No semantic understanding**: It treats words as independent units and doesn't understand synonyms or context
- **Corpus dependency**: TF-IDF scores depend on the entire corpus, so adding or removing documents changes the scores
- **Document length bias**: Can be affected by document length differences (though this is partially addressed by normalization)

For more advanced semantic analysis, techniques like word embeddings or transformer models might be more appropriate.
:::

TF-IDF bridges the gap between simple word counting and more sophisticated text analysis techniques. By weighing words based on both their local importance (in a document) and their global rarity (across the corpus), it helps us discover the vocabulary that truly distinguishes different parts of our text data.
